`r glue::glue('Doublet benchmarking results: {params$dataset}')`

### 2024-08-05 analyses/cell-type-ewings/exploratory_analysis/02-cellassign-tumor-cell-classification.Rmd

```{r}
# function for creating a binary matrix with cell types as columns and marker genes as rows 
build_binary_df <- function(marker_genes_df,
                            gene_col, 
                            cell_type_col) {

  binary_df <- marker_genes_df |>
    dplyr::select({{cell_type_col}}, {{gene_col}}) |> 
    unique() |> 
    tidyr::pivot_wider(
      id_cols = {{gene_col}},
      names_from = {{cell_type_col}},
      values_from = {{cell_type_col}},
      values_fn = length,
      values_fill = 0
    ) |>
    # add a column with no marker genes
    # cell assign will assign cells to "other" when no other cell types are appropriate
    dplyr::mutate(other = 0)

@@

# function to get assigned cell type based on max prediction returned by CellAssign 
get_celltype_assignment <- function(predictions){
  # get individual cell type assignments
  # those with the max prediction 
  celltype_assignments <- predictions |>
    tidyr::pivot_longer(
      !barcode,
      names_to = "celltype",
      values_to = "score"
    ) |>
    dplyr::group_by(barcode) |>
    dplyr::slice_max(score, n = 1) |>
    dplyr::ungroup()

@@

```{r}
tumor_ref_mtx <- build_binary_df(tumor_markers_df, 
                                  gene_col = "ensembl_gene_id",
                                  cell_type_col = "cell_type")

@@

# heatmap of prediction scores 
tumor_predictions |>
  tibble::column_to_rownames("barcode") |> 
  as.matrix() |> 
  pheatmap::pheatmap(show_rownames = FALSE,
                     annotation_row = annotation_df)
```

@@

caret::confusionMatrix(
  table(
    celltype_df$marker_gene_classification, 
    celltype_df$cellassign_annotations)
)
```

@@

```{r}
# heatmap of prediction scores 
filtered_predictions |>
  tibble::column_to_rownames("barcode") |> 
  as.matrix() |> 
  pheatmap::pheatmap(show_rownames = FALSE,
                     annotation_row = annotation_df)
```

@@

```{r}
# build the reference matrix 
visser_ref_mtx <- build_binary_df(all_markers_df, 
                                   gene_col = "ensembl_gene_id",
                                   cell_type_col = "cell_type")

@@

```{r}
# heatmap of prediction scores 
visser_predictions |>
  tibble::column_to_rownames("barcode") |> 
  as.matrix() |> 
  pheatmap::pheatmap(show_rownames = FALSE,
                     annotation_row = annotation_df)
```

@@

caret::confusionMatrix(
  table(
    celltype_df$caret_cellassign_annotations, 
    celltype_df$caret_visser_annotations)
)
```

@@

# get the ensembl id from rowdata for each panglao marker 
panglao_tumor_markers_df <- rowData(sce) |> 
  as.data.frame() |> 
  dplyr::filter(gene_symbol %in% panglao_markers_df$`official gene symbol`) |> 
  dplyr::left_join(panglao_markers_df, by = c("gene_symbol" = "official gene symbol")) |> 
  dplyr::select(gene_symbol, ensembl_gene_id = gene_ids, cell_type = `cell type`) |> 
  # add in the tumor markers 
  dplyr::bind_rows(tumor_markers_df)

```

```{r}
# build the reference matrix 
panglao_tumor_ref_mtx <- build_binary_df(panglao_tumor_markers_df, 
                                          gene_col = "ensembl_gene_id",
                                          cell_type_col = "cell_type")

@@

panglao_tumor_ref_mtx <- panglao_tumor_ref_mtx[-which(rowsums > 1),]

@@

```{r}
# heatmap of prediction scores 
panglao_predictions |>
  tibble::column_to_rownames("barcode") |> 
  as.matrix() |> 
  pheatmap::pheatmap(show_rownames = FALSE,
                     annotation_row = annotation_df)
```

>>>>

```{r}
# function for creating a binary matrix with cell types as columns and marker genes as rows
build_binary_df <- function(marker_genes_df,
                            gene_col,
                            cell_type_col) {
  binary_df <- marker_genes_df |>
    dplyr::select({{ cell_type_col }}, {{ gene_col }}) |>
    unique() |>
    tidyr::pivot_wider(
      id_cols = {{ gene_col }},
      names_from = {{ cell_type_col }},
      values_from = {{ cell_type_col }},
      values_fn = length,
      values_fill = 0
    ) |>
    # add a column with no marker genes
    # cell assign will assign cells to "other" when no other cell types are appropriate
    dplyr::mutate(other = 0)

@@

# function to get assigned cell type based on max prediction returned by CellAssign
get_celltype_assignment <- function(predictions) {
  # get individual cell type assignments
  # those with the max prediction
  celltype_assignments <- predictions |>
    tidyr::pivot_longer(
      !barcode,
      names_to = "celltype",
      values_to = "score"
    ) |>
    dplyr::group_by(barcode) |>
    dplyr::slice_max(score, n = 1) |>
    dplyr::ungroup()

@@

```{r}
tumor_ref_mtx <- build_binary_df(tumor_markers_df,
  gene_col = "ensembl_gene_id",
  cell_type_col = "cell_type"
)

@@

# heatmap of prediction scores
tumor_predictions |>
  tibble::column_to_rownames("barcode") |>
  as.matrix() |>
  pheatmap::pheatmap(
    show_rownames = FALSE,
    annotation_row = annotation_df
  )
```

@@

caret::confusionMatrix(
  table(
    celltype_df$marker_gene_classification,
    celltype_df$cellassign_annotations
  )
)
```

@@

```{r}
# heatmap of prediction scores
filtered_predictions |>
  tibble::column_to_rownames("barcode") |>
  as.matrix() |>
  pheatmap::pheatmap(
    show_rownames = FALSE,
    annotation_row = annotation_df
  )
```

@@

```{r}
# build the reference matrix
visser_ref_mtx <- build_binary_df(all_markers_df,
  gene_col = "ensembl_gene_id",
  cell_type_col = "cell_type"
)

@@

```{r}
# heatmap of prediction scores
visser_predictions |>
  tibble::column_to_rownames("barcode") |>
  as.matrix() |>
  pheatmap::pheatmap(
    show_rownames = FALSE,
    annotation_row = annotation_df
  )
```

@@

caret::confusionMatrix(
  table(
    celltype_df$caret_cellassign_annotations,
    celltype_df$caret_visser_annotations
  )
)
```

@@

# get the ensembl id from rowdata for each panglao marker
panglao_tumor_markers_df <- rowData(sce) |>
  as.data.frame() |>
  dplyr::filter(gene_symbol %in% panglao_markers_df$`official gene symbol`) |>
  dplyr::left_join(panglao_markers_df, by = c("gene_symbol" = "official gene symbol")) |>
  dplyr::select(gene_symbol, ensembl_gene_id = gene_ids, cell_type = `cell type`) |>
  # add in the tumor markers
  dplyr::bind_rows(tumor_markers_df)
```

```{r}
# build the reference matrix
panglao_tumor_ref_mtx <- build_binary_df(panglao_tumor_markers_df,
  gene_col = "ensembl_gene_id",
  cell_type_col = "cell_type"
)

@@

panglao_tumor_ref_mtx <- panglao_tumor_ref_mtx[-which(rowsums > 1), ]

@@

```{r}
# heatmap of prediction scores
panglao_predictions |>
  tibble::column_to_rownames("barcode") |>
  as.matrix() |>
  pheatmap::pheatmap(
    show_rownames = FALSE,
    annotation_row = annotation_df
  )
```

### 2024-08-05 analyses/cell-type-ewings/template_notebooks/cnv-workflow/04-infercnv.Rmd

# png ref files 
png_file <-file.path(params$infercnv_dir, glue::glue("{params$library_id}_infercnv.png"))

@@

# check if marker gene annotations are present 
if(all(is.na(manual_classifications_df$marker_gene_classification))){
  has_marker_gene <- FALSE
  message("No annotations were available using only marker gene expression. 
          Any plots comparing InferCNV to marker gene annotation will be skipped.")
} else {
  has_marker_gene <- TRUE 
}

@@

# don't include the normal cells that were provided as reference in scaling 
cnv_df <- all_cnv_df |>
  # get total number of chromosomes with cnvs per cell and then scale that number 
  dplyr::mutate(
    has_cnv = ifelse(
      stringr::str_detect(subcluster, "reference"), 
      NA_real_, 
      rowSums(across(has_cnv_cols))
    ))
```

```{r}
# plot the total number of CNVs
ggplot(cnv_df, aes(x = UMAP1, y= UMAP2, color = has_cnv)) +
  geom_point(alpha = 0.5, size = 0.5) +
  theme_bw() +
  scale_color_viridis_c()
```

>>>>

# png ref files
png_file <- file.path(params$infercnv_dir, glue::glue("{params$library_id}_infercnv.png"))

@@

# check if marker gene annotations are present
if (all(is.na(manual_classifications_df$marker_gene_classification))) {
  has_marker_gene <- FALSE
  message("No annotations were available using only marker gene expression.
          Any plots comparing InferCNV to marker gene annotation will be skipped.")
} else {
  has_marker_gene <- TRUE
}

@@

# don't include the normal cells that were provided as reference in scaling
cnv_df <- all_cnv_df |>
  # get total number of chromosomes with cnvs per cell and then scale that number
  dplyr::mutate(
    has_cnv = ifelse(
      stringr::str_detect(subcluster, "reference"),
      NA_real_,
      rowSums(across(has_cnv_cols))
    )
  )
```

```{r}
# plot the total number of CNVs
ggplot(cnv_df, aes(x = UMAP1, y = UMAP2, color = has_cnv)) +
  geom_point(alpha = 0.5, size = 0.5) +
  theme_bw() +
  scale_color_viridis_c()
```

### 2024-08-05 analyses/doublet-detection/exploratory-notebooks/03_compare-benchmark-results.Rmd

      scuttle::makePerCellDF(sce, use.dimred = "PCA") |>
        tibble::rownames_to_column(var = "barcodes") |>
        dplyr::select(barcodes,
          ground_truth = ground_truth_doublets,
          PC1 = PCA.1,
          PC2 = PCA.2
        ) |>
        dplyr::left_join(
          scrub_df,
          by = "barcodes"
        ) |>
        dplyr::left_join(
          scdbl_df,
          by = "barcodes"
        )
    }
  ) |>
  purrr::set_names(dataset_names)
```

>>>>

      scuttle::makePerCellDF(sce, use.dimred = "PCA") |>
        tibble::rownames_to_column(var = "barcodes") |>
        dplyr::select(
          barcodes,
          ground_truth = ground_truth_doublets,
          PC1 = PCA.1,
          PC2 = PCA.2
        ) |>
        dplyr::left_join(
          scrub_df,
          by = "barcodes"
        ) |>
        dplyr::left_join(
          scdbl_df,
          by = "barcodes"
        )
    }
  ) |>
  purrr::set_names(dataset_names)
```

### 2024-08-05 analyses/doublet-detection/template-notebooks/02_explore-benchmark-results.Rmd

# grab ground truth and PCA coordinates
sce <- read_rds(sce_file)
sce_df <- scuttle::makePerCellDF(sce, use.dimred = "PCA") |>
  tibble::rownames_to_column(var = "barcodes") |>
  dplyr::select(barcodes,
    ground_truth = ground_truth_doublets,
    PC1 = PCA.1,
    PC2 = PCA.2
  )

>>>>

# grab ground truth and PCA coordinates
sce <- read_rds(sce_file)
sce_df <- scuttle::makePerCellDF(sce, use.dimred = "PCA") |>
  tibble::rownames_to_column(var = "barcodes") |>
  dplyr::select(
    barcodes,
    ground_truth = ground_truth_doublets,
    PC1 = PCA.1,
    PC2 = PCA.2
  )
